Creating the ``r params$package_name`` R package

### 2025-12-30 create-FSHybridPLS.Rmd
before modifying get_rho

---
title: "Creating the ``r params$package_name`` R package"
author: "Your Name"
date: "The Date"
knit: litr::render
output: litr::litr_html_document
params:
  package_name: "FSHybridPLS" # <-- change this to your package name
  package_parent_dir: "." # <-- relative to this file's location
---

@@

We start by specifying the information needed in the DESCRIPTION file of the R package.
 library()
library(fda)
```{r package-setup, message=FALSE, results='hide'}
usethis::create_package(
  path = ".",
  fields = list(
    Package = params$package_name,
    Version = "0.0.0.9000",
    Title = "A Package That Says Hello",
    Description = "This package says hello.  But its actual purpose is to show how an R package can be completely coded in a single R markdown file.",
    Imports = "fda (>= 6.1.3), Matrix, funData, foreach, irlba, refund, R6",
    `Authors@R` = person(
      given = "First",
      family = "Last",
      email = "you@gmail.com",
      role = c("aut", "cre")
      )
  )
)
usethis::use_mit_license(copyright_holder = "F. Last")
```

# helper functions

### is_same_basis
```{r}
#' @export
is_same_basis <- function(input, other) {
  all(
    length(input$functional_list) == length(other$functional_list),
    all(
      mapply(function(fd1, fd2) fda::is.eqbasis(fd1$basis, fd2$basis),
             input$functional_list, other$functional_list)
    )
  )
}
```

### compute_gram_matrix
```{r}
#' Compute the Gram matrix for a basis object
#'
#' @param basis A basis object of class `basisfd` from the `fda` package
#' @return A square matrix where entry (i,j) is the inner product of basis functions i and j
#' @export
compute_gram_matrix <- function(basis) {
  stopifnot(inherits(basis, "basisfd"))
  inprod(basis, basis)  # computes ∫ b_i(t) b_j(t) dt
}

@@

## helpers for functional data objects

@@

```{r}
#' Replicate a list of single-sample fd objects multiple times
#'
#' This function broadcasts a list of single-sample functional predictors
#' by replicating their coefficient columns.
#'
#' @param fd_list A list of `fd` objects, each representing a single sample.
#' @param n The number of replications desired.
#'
#' @return A list of `fd` objects, each with `n` replications.
#' @export
rep_fd <- function(fd_list, n) {
  if (!is.list(fd_list) || any(!vapply(fd_list, inherits, logical(1), "fd"))) {
    stop("Input must be a list of 'fd' objects.")
  }

  lapply(fd_list, function(fd_obj) {
    coef_mat <- fd_obj$coefs
    if (is.null(dim(coef_mat)) || ncol(coef_mat) != 1) {
      stop("Each fd object in the list must have one column of coefficients.")
    }

    new_coefs <- matrix(rep(coef_mat, n), nrow = nrow(coef_mat), ncol = n)
    fd(coef = new_coefs, basisobj = fd_obj$basis)
  })
}
```

| Case | Description                            | Input Type               | Expected Output                        | Notes                                            |
|------|----------------------------------------|--------------------------|----------------------------------------|--------------------------------------------------|
| 1    | Correct replication                    | List of single-sample fd | List of `fd` objects with `n` columns | All columns identical to original coefficients   |
| 2    | Each fd has one sample                 | List of fd               | No error                               | Verifies structure and content                  |
| 3    | One fd has multiple samples            | fd with >1 columns       | Error                                   | Should raise informative error                  |
| 4    | Non-fd object in list                  | List of non-`fd` items   | Error                                   | Validates class-checking robustness             |

# Class predictor_hybrid
- Let $\{X^{(k)}\}_{k=1, \ldots, K}$ be a collection of random functions defined on unit hypercube domains $\mathcal{T}_k := [0,1]^{d_k}$ ($d_k \in \mathbb{N})$; i.e., $X^{(k)}: \mathcal{T}_k \rightarrow \mathbb{R}$.

- Assume that each $X^{(k)}$ is in $L^2(\mathcal{T}_k)$, a Hilbert space of square integrable functions with respect to Lebesgue measure $dt_k$ on $\mathcal{T}_k$.

- Write  $X=(X^{(1)}, \ldots, X^{(K)})$ as a multivariate functional object that belongs to $\mathcal{F} = L^2(\mathcal{T}_1) \times \cdots \times L^2(\mathcal{T}_K)$---a cartesian product of individual $L^2(\mathcal{T}_k)$ spaces.

- We can also express the functional object $X$ evaluated on the multi-dimensional argument $\mathbf{t} = (t_1, \ldots, t_k)^\top \in \mathcal{T} = \mathcal{T}_1 \times \cdots \times \mathcal{T}_K$ as a $K$-dimensional vector $X(\mathbf{t})=(X^{(1)}(t_1), \ldots, X^{(K)}(t_K))^\top$.

- **Our strategy is to formulate a hybrid random object**, $\mathbf{W} = (X, \mathbf{Z})$, which combines $X$ and $p$-dimensional scalar covariate $\mathbf{Z}$ into an ordered pair belonging to $\mathcal{H} = \mathcal{F} \times \mathbb{R}^p$.

- An alternative notation for the hybrid object can be obtained by evaluating its functional part on $\mathbf{t}$ and expressing it as a $(K+p)$-dimensional vector: $\mathbf{W}[\mathbf{t}] = (X(\mathbf{t}), \mathbf{Z})^\top$, with $X(\mathbf{t}) = (X^{(1)}(t_1), \ldots, X^{(K)}(t_K))^\top \in \mathbb{R}^K$ and $\mathbf{Z} = (Z_1, \ldots, Z_p)^\top  \in \mathbb{R}^p$.

- The hybrid object forms a Hilbert space, characterized by well-defined addition, scalar multiplication, and inner product operations.

- We define an S3 object `predictor_hybrid` that represent this structure, and implement Hilbert space operations for this class, and our algorithm is built on these foundational operations.

The class definition is a simple list that contains a matrix and a collection of \code{fd} objects from the \code{fda} package, along with the necessary metadata.
- gram matrices are saved as a list.
```{r classdef:predictor_hybrid}
#' Create a predictor_hybrid object (S3 version, automatic basis size and Gram matrix)
#' Constructs a hybrid predictor object that stores both scalar and functional predictors,
#' and automatically computes Gram matrices (inner products of basis functions).
#'
#' @param Z A numeric matrix of dimension \code{n_sample × n_scalar} representing scalar predictors.
#' @param functional_list A list of functional predictors (e.g., \code{fd} objects from the \code{fda} package).
#'
#' @return An object of class \code{predictor_hybrid}, containing scalar and functional data with Gram matrices.
#' @export
predictor_hybrid <- function(Z, functional_list, eval_point) {
  stopifnot(is.matrix(Z), is.numeric(Z))
  stopifnot(is.list(functional_list))

@@

  gram_list <- gram_deriv2_list <- vector("list", n_functional)
  n_basis_list <- numeric(n_functional)

  for (i in seq_len(n_functional)) {
    fd_i <- functional_list[[i]]
    stopifnot(inherits(fd_i, "fd"))

    n_fd_sample <- ncol(coef(fd_i))
    if (n_fd_sample != n_sample) {
      stop(sprintf("Functional predictor %d has %d replicates, but Z has %d rows.",
                   i, n_fd_sample, n_sample))
    }

    basis_i <- fd_i$basis
    gram_list[[i]] <- compute_gram_matrix(basis_i)
    gram_deriv2_list[[i]] <- fda::getbasispenalty(basis_i) # New

@@

  structure(
    list(
      Z = Z,
      functional_list = functional_list,
      eval_point = eval_point,
      gram_list = gram_list,
      gram_deriv2_list = gram_deriv2_list, # Store the new list
      n_basis_list = n_basis_list,
      n_sample = n_sample,
      n_functional = n_functional,
      n_scalar = n_scalar
    ),
    class = "predictor_hybrid"
  )
}
```

### predictor_hybrid_from_coef
An alternative constructor for a predictor_hybrid object based on a flat numeric coefficient vector. It reconstructs both the functional and scalar components for a single-sample hybrid predictor object. The function reads the basis information from the template `format` object.

- Inputs
  - **`format`**: A `predictor_hybrid` object containing:
    - `functional_list`: List of `fd` objects with basis info
    - `gram_list`: List of Gram matrices
    - `n_scalar`: Number of scalar predictors
  - **`coef`**: Numeric vector of length `K * M + d`, where:
    - `K` = number of functional predictors  
    - `M` = number of basis functions (assumed same across predictors)  
    - `d` = number of scalar predictors

- Output
  - A `predictor_hybrid` object with:
    - Functional predictors rebuilt from the first `K * M` coefficients
    - Scalar predictor `Z` from the remaining `d` coefficients
    - `n_sample` set to 1

```{r}
#' Construct a Single-Sample Predictor Hybrid Object from Coefficients
#' Reconstructs a \code{predictor_hybrid} object representing one observation, using a numeric
#' coefficient vector. This alternative constructor maps the coefficients back into their functional
#' and scalar predictor representations based on the structure of a template \code{predictor_hybrid} object.
#'
#' @param format A \code{predictor_hybrid} object that provides the structure and basis information.
#' @param coef A numeric vector containing coefficients for both functional and scalar predictors.
#'
#' @return A \code{predictor_hybrid} object with updated \code{functional_list}, \code{Z}, and \code{n_sample = 1}.
#'
#'
#' @export
predictor_hybrid_from_coef <- function(format, coef){
  M <- format$n_basis_list[1] # number of basis functions
  K <- format$n_functional # number of functional predictors
  for (ii in 1:K){
    format$functional_list[[ii]] <- fd(
      coef = as.matrix(coef[((ii-1)*M + 1):(ii*M)]),
      basisobj = format$functional_list[[ii]]$basis
    )
  }
  format$Z <- t(as.matrix(coef[(K * M + 1):length(coef)]))
  format$n_sample <- 1
  return(format)
}
```

@@

### add.predictor_hybrid
Performs element-wise addition of two `predictor_hybrid` objects.  Functional predictors are combined using `plus.fd()` and `times.fd()` from the `fda` package.

- Usage

```r
add.predictor_hybrid(input, other, alpha = 1)
```

- Arguments:

@@

- Value:
Returns a new `predictor_hybrid` object representing the result of the addition.

- Details: 
  - This function assumes both objects have the same number and structure of functional and scalar predictors.  
  - Functional parts are scaled by `alpha` using `times.fd()` and then summed using `plus.fd()`.  
  - Scalar predictors are added using standard matrix addition.

```{r}
add.predictor_hybrid <- function(xi_1, xi_2, alpha = 1) {
  # Safe access to is.eqbasis()
  is_eqbasis <- getFromNamespace("is.eqbasis", "fda")

  for (i in seq_len(xi_1$n_functional)) {
    if (!is_eqbasis(xi_1$functional_list[[i]]$basis, xi_2$functional_list[[i]]$basis)) {
      stop("Functional predictors must have the same basis.")
    }
  }

@@

  # Swap so that broadcasting always applies to xi_2
  if (xi_1$n_sample == 1 && xi_2$n_sample > 1) {
    tmp <- xi_1
    xi_1 <- xi_2
    xi_2 <- tmp
  }

@@

  # Replicate fd and Z if needed
  if (n2 == 1) {
    f2 <- rep_fd(f2, n1)
    Z2 <- matrix(rep(c(Z2), n1), nrow = n1, byrow = TRUE)
  }

  # Combine functional predictors
  new_functional_list <- Map(
    function(f1, f2) plus.fd(f1, times.fd(alpha, f2)),
    f1,
    f2
  )

  # Compute scalar inner products
  inprod_scalar <- rowSums(Z1 * Z2) 
  # Combine scalar predictors
  new_Z <- Z1 + alpha * Z2

  # Construct new object
  predictor_hybrid(Z = new_Z, functional_list = new_functional_list)
}
```

### subtr.predictor_hybrid
Performs element-wise subtraction of two `predictor_hybrid` objects.   Internally uses `add.predictor_hybrid(input, other, alpha = -1)` to compute the result by negating the second operand.

- Usage

```r
subtr.predictor_hybrid(input, other, alpha = 1)
```

- Arguments:

- Value:  
Returns a new `predictor_hybrid` object representing the result of the subtraction.

- Details:  
  - This function assumes both objects have the same number and structure of functional and scalar predictors.  
  - It performs subtraction by internally calling `add.predictor_hybrid()` with `-alpha`.  
  - Functional parts are scaled using `times.fd()` and subtracted via `plus.fd()` with a negated factor.  
  - Scalar predictors are subtracted using standard matrix arithmetic.

```{r}
#' Subtract two predictor_hybrid objects
#'
#' Performs element-wise subtraction of two `predictor_hybrid` objects.
#' Internally uses `add.predictor_hybrid(input, other, alpha = -1)`.
#'
#' @param input A `predictor_hybrid` object.
#' @param other Another `predictor_hybrid` object to subtract.
#' @param alpha A scalar multiplier applied to `other` before subtraction (default is 1).
#'
#' @return A new `predictor_hybrid` object representing the result of subtraction.
#' @export
subtr.predictor_hybrid <- function(input, other, alpha = 1) {
  add.predictor_hybrid(input, other, alpha = -alpha)
}

```
We skip the unit test for this function.

### scalar_mul.predictor_hybrid
Multiplies all components of a `predictor_hybrid` object by a scalar. Functional components are scaled using `times.fd()` from the `fda` package, and scalar predictors are multiplied directly using matrix operations.

- Usage

```r
scalar_mul.predictor_hybrid(input, scalar)
```

- Arguments:

- Value:  
Returns a new `predictor_hybrid` object with all components scaled by `scalar`.

- Details:  
  - Functional predictors are scaled using `times.fd(scalar, fd_obj)` for each element.  
  - Scalar predictors (the matrix `Z`) are scaled elementwise using matrix multiplication.

  # Reconstruct hybrid object
  predictor_hybrid(
    Z = new_Z,
    functional_list = new_functional_list
  )
}
```

>>>>

---
title: "Creating the ``r params$package_name`` R package"
author: "Jongmin Mun"
date: "The Date"
knit: litr::render
output: litr::litr_html_document
params:
  package_name: "FSHybridPLS" # <-- change this to your package name
  package_parent_dir: "." # <-- relative to this file's location
---

@@

We start by specifying the information needed in the DESCRIPTION file of the R package.

```{r package-setup, message=FALSE, results='hide'}
usethis::create_package(
  path = ".",
  fields = list(
    Package = params$package_name,
    Version = "0.0.0.9000",
Title = "Hybrid Penalized Partial Least Squares for Functional and Scalar Data",
    Description = "Implements a Penalized Partial Least Squares (PLS) regression framework designed for hybrid predictors that combine both functional (infinite-dimensional) and scalar (finite-dimensional) covariates. The algorithm maximizes the covariance between the hybrid predictors and a scalar response variable while incorporating a roughness penalty to ensure the smoothness of the estimated functional coefficient curves. The package includes tools for data simulation using matrix-normal distributions, two-step data normalization, and cross-validated model fitting.",
    Imports = "fda (>= 6.1.3), Matrix, funData, foreach, irlba, refund, mvnfast, R6",
    `Authors@R` = person(
      given = "First",
      family = "Last",
      email = "you@gmail.com",
      role = c("aut", "cre")
      )
  )
)
usethis::use_mit_license(copyright_holder = "F. Last")
```

#### Table of Contents
- [Helper Functions](#helper)

# Helper Functions {#helper}

### compute_gram_matrix

**Description:**
Calculates the Gram matrix for a given functional basis. The Gram matrix $G$ is a symmetric positive semi-definite matrix containing the inner products of the basis functions. Specifically, the entry $(i, j)$ corresponds to the integral of the product of the $i$-th and $j$-th basis functions over the domain:
$$G_{ij} = \langle b_i, b_j \rangle = \int b_i(t) b_j(t) \, dt$$

**Inputs:**

- `basis`: A basis object of class `basisfd` (typically created using the `fda` package).

**Output:**

- Returns a square numeric matrix where the entry at row $i$ and column $j$ is the inner product of basis functions $i$ and $j$.

```{r}
#' Compute the Gram matrix for a basis object
#'
#' This function calculates the inner product of a basis with itself to form
#' the Gram matrix.
#'
#' @param basis A basis object of class `basisfd` from the `fda` package.
#' @return A square matrix where entry (i,j) is the inner product of basis functions i and j.
#' @export
compute_gram_matrix <- function(basis) {
  # Ensure the input is a valid basis object from the fda package
  stopifnot(inherits(basis, "basisfd"))

  # Compute the inner product of the basis with itself
  # This calculates the integral of b_i(t) * b_j(t) using the metric defined on the basis
  fda::inprod(basis, basis) 
}
```

### is_same_basis

**Description:**
Checks if two composite functional data objects (represented as lists containing functional data) share the exact same basis functions. It verifies that both lists have the same length and that the basis for every corresponding element is identical.

**Inputs:**

- `input`: A list containing functional data objects (expected to have a `$functional_list` structure).
- `other`: Another list containing functional data objects to compare against.

**Output:**

- Returns `TRUE` if both lists have the same length and every corresponding basis is equivalent; otherwise returns `FALSE`.

```{r}
#' Check if two functional data lists share the same basis
#'
#' @description
#' Checks if two composite functional data objects share the exact same basis functions
#' by comparing their internal list structures.
#'
#' @param input A list containing functional data objects (expected to have a `$functional_list` structure).
#' @param other Another list containing functional data objects to compare against.
#' @return `TRUE` if both lists have the same length and every corresponding basis is equivalent; otherwise `FALSE`.
#' @export
is_same_basis <- function(input, other) {
  all(
    # Check if both lists have the same number of functional elements
    length(input$functional_list) == length(other$functional_list),

    # Check if the basis for each corresponding element is identical
    all(
      mapply(function(fd1, fd2) {
        # Use fda package function to check basis equality
        fda::is.eqbasis(fd1$basis, fd2$basis)
      }, input$functional_list, other$functional_list)
    )
  )
}
```

### are_all_gram_matrices_identical

**Description:**
Iterates through a list of matrices to verify if they are all numerically identical to the first matrix in the list. This function handles floating-point inaccuracies by using a tolerance parameter.

**Inputs:**

- `gram_list`: A list containing matrix objects to be compared.
- `tolerance` (Optional): A numeric value defining the tolerance for floating-point comparison. Defaults to the square root of machine epsilon.

**Output:**

- Returns `TRUE` if the list is empty, has one element, or if all matrices are identical to the first. Returns `FALSE` if any matrix differs.

@@

@@

**Description:**
Takes a list of functional data objects (where each object represents a single sample/curve) and broadcasts them by replicating their coefficients. This creates new functional objects containing $n$ identical copies of the original curve.

**Inputs:**

- `fd_list`: A list of `fd` objects (functional data objects). Each object must have exactly one column of coefficients (a single sample).
- `n`: An integer specifying the number of replications desired.

**Output:**

- Returns a list of `fd` objects, where each object now contains $n$ replications of the original coefficients.

```{r}
#' Replicate a list of single-sample fd objects multiple times
#'
#' This function broadcasts a list of single-sample functional predictors
#' by replicating their coefficient columns.
#'
#' @param fd_list A list of `fd` objects, each representing a single sample.
#' @param n The number of replications desired.
#'
#' @return A list of `fd` objects, each with `n` replications.
#' @export
rep_fd <- function(fd_list, n) {
  # Validate input: must be a list of 'fd' objects
  if (!is.list(fd_list) || any(!vapply(fd_list, inherits, logical(1), "fd"))) {
    stop("Input must be a list of 'fd' objects.")
  }

  lapply(fd_list, function(fd_obj) {
    coef_mat <- fd_obj$coefs

    # Ensure the object represents a single sample (one column of coefficients)
    if (is.null(dim(coef_mat)) || ncol(coef_mat) != 1) {
      stop("Each fd object in the list must have one column of coefficients.")
    }

    # Create a new coefficient matrix by repeating the column 'n' times
    new_coefs <- matrix(rep(coef_mat, n), nrow = nrow(coef_mat), ncol = n)

    # Return a new fd object with the replicated coefficients and original basis
    fd(coef = new_coefs, basisobj = fd_obj$basis)
  })
}
```

# Hybrid predictor class

## Class definition and constructor
### predictor_hybrid

**Description:**
The `predictor_hybrid` class represents a hybrid random object $\mathbf{W} = (X, \mathbf{Z})$ that combines functional and scalar covariates into a unified Hilbert space structure.

- **Functional Part:** Let $\{X^{(k)}\}_{k=1, \ldots, K}$ be a collection of random functions defined on unit interval $\mathcal{T}_k := [0,1]$. Each $X^{(k)}$ belongs to $L^2([0,1])$, a Hilbert space of square-integrable functions. The multivariate functional object $X$ resides in the cartesian product space $\mathcal{F} = L^2([0,1]) \times \cdots \times L^2([0,1])$.
- **Hybrid Object:** We define the hybrid object $\mathbf{W} = (X, \mathbf{Z})$, where $\mathbf{Z}$ is a $p$-dimensional scalar covariate vector. This object belongs to the product space $\mathcal{H} = \mathcal{F} \times \mathbb{R}^p$.
- **Vector Notation:** The object can be evaluated at a multi-dimensional argument $\mathbf{t}$ as a $(K+p)$-dimensional vector: $\mathbf{W}[\mathbf{t}] = (X(\mathbf{t}), \mathbf{Z})^\top$.
- **Implementation:** The class is implemented as an S3 object containing a matrix for scalar parts, a list of `fd` objects for functional parts, and pre-computed Gram matrices to facilitate Hilbert space operations.

**Inputs:**

- `Z`: A numeric matrix of dimension $n\_sample \times n\_scalar$ representing the scalar predictors.
- `functional_list`: A list of functional predictors (typically `fd` objects from the `fda` package).
- `eval_point`: A specification for evaluation points (stored as metadata).

**Output:**

- Returns an object of class `predictor_hybrid`, which includes the original data, pre-computed Gram matrices, and roughness penalty matrices.

```{r}
#' Create a predictor_hybrid object (S3 version, automatic basis size and Gram matrix)
#'
#' Constructs a hybrid predictor object that stores both scalar and functional predictors,
#' and automatically computes Gram matrices (inner products of basis functions).
#'
#' @param Z A numeric matrix of dimension \code{n_sample x n_scalar} representing scalar predictors.
#' @param functional_list A list of functional predictors (e.g., \code{fd} objects from the \code{fda} package).
#' @param eval_point Points at which the functions are evaluated (stored as metadata).
#'
#' @return An object of class \code{predictor_hybrid}, containing scalar and functional data with Gram matrices.
#' @export
predictor_hybrid <- function(Z, functional_list, eval_point) {
  # Validate scalar input
  stopifnot(is.matrix(Z), is.numeric(Z))

  # Validate functional input
  stopifnot(is.list(functional_list))

@@

  # Initialize lists for storing matrices and metadata
  gram_list <- gram_deriv2_list <- vector("list", n_functional)
  n_basis_list <- numeric(n_functional)

  # Iterate through each functional predictor to compute necessary matrices
  for (i in seq_len(n_functional)) {
    fd_i <- functional_list[[i]]
    stopifnot(inherits(fd_i, "fd"))

    # Ensure the number of functional samples matches the scalar samples
    n_fd_sample <- ncol(coef(fd_i))
    if (n_fd_sample != n_sample) {
      stop(sprintf("Functional predictor %d has %d replicates, but Z has %d rows.",
                   i, n_fd_sample, n_sample))
    }

    # Extract basis and compute inner product matrices
    basis_i <- fd_i$basis
    gram_list[[i]] <- compute_gram_matrix(basis_i)
    gram_deriv2_list[[i]] <- fda::getbasispenalty(basis_i) # Compute roughness penalty matrix

@@

  # Construct the S3 object
  structure(
    list(
      Z = Z,
      functional_list = functional_list,
      eval_point = eval_point,
      gram_list = gram_list,
      gram_deriv2_list = gram_deriv2_list, # Store the penalty matrices
      n_basis_list = n_basis_list,
      n_sample = n_sample,
      n_functional = n_functional,
      n_scalar = n_scalar
    ),
    class = "predictor_hybrid"
  )
}
```

### predictor_hybrid_from_coef

**Description:**
An alternative constructor that creates a single-sample `predictor_hybrid` object from a flat numeric coefficient vector. It reconstructs both the functional and scalar components by mapping the coefficients back onto the basis functions defined in a template `format` object.

**Important:** This function strictly assumes that all functional predictors in the `format` object share the same number of basis functions (). It verifies this assumption before proceeding.

**Inputs:**

- `format`: A `predictor_hybrid` object serving as a template. It must contain the necessary basis information (`functional_list`) and structural metadata (`n_scalar`, `n_functional`).
- `coef`: A numeric vector of length , where:
-  is the number of functional predictors.
-  is the number of basis functions (strictly enforced to be identical across all functional predictors).
-  is the number of scalar predictors.

**Output:**

- Returns a `predictor_hybrid` object representing a single sample (`n_sample = 1`), containing:
- Functional predictors rebuilt from the first  coefficients.
- A scalar predictor matrix `Z` derived from the remaining  coefficients.

```{r}
#' Construct a Single-Sample Predictor Hybrid Object from Coefficients
#'
#' Reconstructs a \code{predictor_hybrid} object representing one observation, using a numeric
#' coefficient vector. This alternative constructor maps the coefficients back into their functional
#' and scalar predictor representations based on the structure of a template \code{predictor_hybrid} object.
#'
#' @param format A \code{predictor_hybrid} object that provides the structure and basis information.
#' @param coef A numeric vector containing coefficients for both functional and scalar predictors.
#'
#' @return A \code{predictor_hybrid} object with updated \code{functional_list}, \code{Z}, and \code{n_sample = 1}.
#' @export
predictor_hybrid_from_coef <- function(format, coef) {
  # Extract metadata from the template
  basis_counts <- format$n_basis_list

  # Check if all functional predictors have the same number of basis functions
  if (length(unique(basis_counts)) != 1) {
    stop("All functional predictors must have the same number of basis functions (M).")
  }

  M <- basis_counts[1]  # Number of basis functions (identical across predictors)
  K <- format$n_functional # Number of functional predictors

  # Reconstruct each functional predictor
  for (ii in 1:K) {
    # Calculate indices for the current functional predictor in the flat vector
    start_idx <- (ii - 1) * M + 1
    end_idx   <- ii * M

    # Create a new fd object using the sliced coefficients and original basis
    format$functional_list[[ii]] <- fd(
      coef = as.matrix(coef[start_idx:end_idx]),
      basisobj = format$functional_list[[ii]]$basis
    )
  }

  # Reconstruct the scalar predictor part (Z)
  # The remaining coefficients after K*M belong to the scalar predictors
  scalar_start_idx <- K * M + 1
  format$Z <- t(as.matrix(coef[scalar_start_idx:length(coef)]))

  # Update sample count to 1 as this constructor creates a single observation
  format$n_sample <- 1

  return(format)
}
```

@@

### add.predictor_hybrid

**Description:**
Performs element-wise addition of two `predictor_hybrid` objects. Functional predictors are combined using `plus.fd()` and `times.fd()` from the `fda` package.

**Usage:**
```r
add.predictor_hybrid(input, other, alpha = 1)
```

**Arguments:**

@@

**Value:**

- Returns a new `predictor_hybrid` object representing the result of the addition.

**Details:**

- This function assumes both objects have the same number and structure of functional and scalar predictors.
- Functional parts are scaled by `alpha` using `times.fd()` and then summed using `plus.fd()`.
- Scalar predictors are added using standard matrix addition.
- Supports broadcasting if one object represents a single sample (`n_sample = 1`) and the other represents multiple samples.

```{r}
#' Add two predictor_hybrid objects
#'
#' Performs element-wise addition of two \code{predictor_hybrid} objects. 
#' Supports broadcasting if one object is a single sample.
#'
#' @param xi_1 A \code{predictor_hybrid} object.
#' @param xi_2 Another \code{predictor_hybrid} object to be added.
#' @param alpha A scalar multiplier applied to \code{xi_2} before addition (default is 1).
#'
#' @return A new \code{predictor_hybrid} object representing the result of the addition.
#' @export
add.predictor_hybrid <- function(xi_1, xi_2, alpha = 1) {
  # Safe access to is.eqbasis() from the fda namespace
  is_eqbasis <- getFromNamespace("is.eqbasis", "fda")

  # Structural checks: Number of functional and scalar predictors must match
  if (xi_1$n_functional != xi_2$n_functional) {
    stop("Mismatch in number of functional predictors.")
  }
  if (xi_1$n_scalar != xi_2$n_scalar) {
    stop("Mismatch in number of scalar predictors.")
  }

  # Basis checks: Ensure corresponding functional predictors share the same basis
  for (i in seq_len(xi_1$n_functional)) {
    if (!is_eqbasis(xi_1$functional_list[[i]]$basis, xi_2$functional_list[[i]]$basis)) {
      stop("Functional predictors must have the same basis.")
    }
  }

  # Broadcasting Logic:
  # Ensure xi_1 is always the larger object (or equal) to simplify broadcasting
  if (xi_1$n_sample == 1 && xi_2$n_sample > 1) {
    tmp <- xi_1
    xi_1 <- xi_2
    xi_2 <- tmp
  }

  # Validate sample compatibility
  if (!(n1 == n2 || n2 == 1)) {
    stop("Sample sizes are incompatible for broadcasting.")
  }

  # Replicate fd and Z if broadcasting is needed (n2 == 1)
  if (n2 == 1 && n1 > 1) {
    f2 <- rep_fd(f2, n1)
    Z2 <- matrix(rep(c(Z2), n1), nrow = n1, byrow = TRUE)
  }

  # Combine functional predictors: f1 + alpha * f2
  new_functional_list <- Map(
    function(fd1, fd2) plus.fd(fd1, times.fd(alpha, fd2)),
    f1,
    f2
  )

  # Combine scalar predictors: Z1 + alpha * Z2
  new_Z <- Z1 + alpha * Z2

  # Construct and return the new predictor_hybrid object
  # Note: eval_point is carried over from the first object
  predictor_hybrid(Z = new_Z, functional_list = new_functional_list, eval_point = xi_1$eval_point)
}
```

### subtr.predictor_hybrid

**Description:**
Performs element-wise subtraction of two `predictor_hybrid` objects. Internally uses `add.predictor_hybrid` to compute the result by negating the second operand.

**Usage:**
```r
subtr.predictor_hybrid(input, other, alpha = 1)
```

**Arguments:**

**Value:**

- Returns a new `predictor_hybrid` object representing the result of the subtraction.

**Details:**

- This function assumes both objects have the same number and structure of functional and scalar predictors.
- It performs subtraction by internally calling `add.predictor_hybrid()` with `-alpha`.
- Functional parts are scaled using `times.fd()` and subtracted via `plus.fd()` with a negated factor.
- Scalar predictors are subtracted using standard matrix arithmetic.

```{r}
#' Subtract two predictor_hybrid objects
#'
#' Performs element-wise subtraction of two `predictor_hybrid` objects.
#' Internally uses `add.predictor_hybrid(input, other, alpha = -1)`.
#'
#' @param input A `predictor_hybrid` object.
#' @param other Another `predictor_hybrid` object to subtract.
#' @param alpha A scalar multiplier applied to `other` before subtraction (default is 1).
#'
#' @return A new `predictor_hybrid` object representing the result of subtraction.
#' @export
subtr.predictor_hybrid <- function(input, other, alpha = 1) {
  add.predictor_hybrid(input, other, alpha = -alpha)
}
```

### scalar_mul.predictor_hybrid

**Description:**
Multiplies all components of a `predictor_hybrid` object by a scalar. Functional components are scaled using `times.fd()` from the `fda` package, and scalar predictors are multiplied directly using matrix operations.

**Usage:**
```r
scalar_mul.predictor_hybrid(input, scalar)

**Arguments:**

**Value:**

- Returns a new `predictor_hybrid` object with all components scaled by `scalar`.

**Details:**

- Functional predictors are scaled using `times.fd(scalar, fd_obj)` for each element.
- Scalar predictors (the matrix `Z`) are scaled elementwise using matrix multiplication.

  # Reconstruct hybrid object
  # Passing eval_point from the original input to maintain metadata
  predictor_hybrid(
    Z = new_Z,
    functional_list = new_functional_list,
    eval_point = input$eval_point
  )
}
```

### inprod.predictor_hybrid

**Description:**
Computes the inner product between two `predictor_hybrid` objects. The inner product in the hybrid space  is defined as the sum of the inner products of the functional components in  and the scalar components in :

**Inputs:**

- `xi_1`: A `predictor_hybrid` object.
- `xi_2`: (Optional) Another `predictor_hybrid` object. Defaults to `xi_1` (computes norm squared if `xi_1` is a single sample).

**Output:**

- A numeric vector of inner products. If both inputs are single samples, returns a scalar.

**Details:**

- Supports broadcasting: if one input has a single observation (`n_sample = 1`) and the other has multiple, the single observation is broadcast across the others.
- Uses `fda::inprod` for functional integration and matrix multiplication for scalar parts.
```{r}
#' Inner product between two predictor_hybrid objects (with broadcasting)
#'
#' Computes the inner product between two `predictor_hybrid` objects,
#' including both functional and scalar components. Supports broadcasting
#' when one of the inputs has a single observation.
#'
#' @param xi_1 A `predictor_hybrid` object.
#' @param xi_2 Another `predictor_hybrid` object. If missing, defaults to `xi_1`.
#'
#' @return A numeric vector of inner products, or a scalar if both inputs contain a single observation.
#' @export
inprod.predictor_hybrid <- function(xi_1, xi_2 = NULL) {
  # Handle self-inner product
  if (is.null(xi_2)) xi_2 <- xi_1

@@

  # Swap so that broadcasting always applies to xi_2 (the smaller one)
  if (xi_1$n_sample == 1 && xi_2$n_sample > 1) {
    tmp <- xi_1
    xi_1 <- xi_2
    xi_2 <- tmp
  }

@@

  # Replicate fd and Z if needed for broadcasting
  if (n2 == 1 && n1 > 1) {
    f2 <- rep_fd(f2, n1)
    Z2 <- matrix(rep(c(Z2), n1), nrow = n1, byrow = TRUE)
  }

  # Compute functional inner products
  # Iterates over samples (i) and sums the inner products of all functional predictors (j)
  inprod_functional <- vapply(seq_len(n1), function(i) {
    sum(vapply(seq_along(f1), function(j) {
      fda::inprod(f1[[j]][i], f2[[j]][i])
    }, numeric(1)))
  }, numeric(1))

  # Compute scalar inner products using row-wise summation
  inprod_scalar <- rowSums(Z1 * Z2)

  # Combine functional and scalar results
  result <- inprod_functional + inprod_scalar

  # Return scalar if both inputs were single samples
  if (n1 == 1 && n2 == 1) as.numeric(result) else result
}
```

### inprod_pen.predictor_hybrid

**Description:**
Computes the penalized inner product between two `predictor_hybrid` objects. This adds a roughness penalty term to the standard inner product, typically used for regularization.

**Inputs:**

- `xi_1`: A `predictor_hybrid` object.
- `xi_2`: (Optional) Another `predictor_hybrid` object.
- `lambda`: A numeric vector of length  (number of functional predictors) specifying the smoothing parameter for each function.

**Output:**

- A numeric vector (or scalar) representing the penalized inner product.
```{r}
#' Penalized Inner product between two predictor_hybrid objects
#'
#' Computes the inner product between two `predictor_hybrid` objects,
#' adding a roughness penalty term based on the second derivative.
#'
#' @param xi_1 A `predictor_hybrid` object.
#' @param xi_2 Another `predictor_hybrid` object. If missing, defaults to `xi_1`.
#' @param lambda A numeric vector of smoothing parameters (one for each functional predictor).
#'
#' @return A numeric vector of inner products.
#' @export
inprod_pen.predictor_hybrid <- function(xi_1, xi_2 = NULL, lambda) {

  # Compute the standard inner product first
  base_inprod <- inprod.predictor_hybrid(xi_1, xi_2)

  # Handle self-reference for xi_2
  if (is.null(xi_2)) xi_2 <- xi_1

  f1 <- xi_1$functional_list
  f2 <- xi_2$functional_list

  # Add penalty terms for each functional predictor
  # The penalty is lambda * <D^2 f1, D^2 f2>
  for (j in seq_len(xi_1$n_functional)) {
    base_inprod <- base_inprod + lambda[j] * fda::inprod(
      fdobj1 = f1[[j]],
      fdobj2 = f2[[j]],
      Lfdobj1 = 2, # Second derivative
      Lfdobj2 = 2  # Second derivative
    )
  }

  return(as.numeric(base_inprod))
}
```
